---
title: "R Notebook"
output: html_document
---

* config() is used to manage your computer and user specific settings, such as location of datadir.
* Configuration is loaded from a config-file (a yaml file), but you can also use config() to set a parameter in code.

* config_file() can be used to get or set the location of the configuration file.
* Configuration file is identified using the following order (the first hit is used):
	* A previously set value in code using: config_file('/path/to/your/config.yml')
	* The value of R_CONFIG_FILE environment variable.
	* One of the following filenames, each searched in the current folder and parent folders:
		* config.[packagename].yml    #e.g., config.immunotar.yml. this is only used if config is running within an installed package.
	  * config.[yourcomputername].yml
  	* config.yml
	* One of the following filenames, searched in current user's home folder:
		* rconfig.[packagename].yml
	  * rconfig.[yourcomputername].yml
  	* rconfig.yml

* config file is loaded as a named list. Any of the fields can be treated as a subconfig.
* 'default' subconfig, if available is always loaded first.
* A subconfig can be selected with config_subconfig('yoursubconfigname') or using the environment variable R_CONFIG_SUBCONFIG.


```{r}
config_file() #this will print the active config file.
config_subconfig() #this will print the active subconfig

# Basic usage: just call config(varname)
config('datadir')  #this may be null. use sys_datadir() when you need it.
config('biodbdir')

# To change a setting from code, use, e.g.,
config('myvariable','myvalue')
# You can later retrieve that setting:
config('myvariable')

# If you make a change to the config.yml file and want the changes to be reloaded, call config_reset(). The settings that are changed manually (ie., from code) will not be reset; if you want to reset them as well, call config_reset(resetcoded=TRUE).
config_reset()

```


```{r}
source('util.r')
source('config.r')

config('config.dbg',T);
config_reset()

config_file()
config_subconfig()
config('datadir')

catf('zoz.config.coded: \n%s\n', str_indentlines(yaml::as.yaml(.GlobalEnv$zoz.config.coded)))
catf('zoz.config: \n%s\n',str_indentlines(yaml::as.yaml(.GlobalEnv$zoz.config)))

```
```{r}

```
 