---
title: "Optimize Immunotar Parameters using Pancancer Dataset"
date: "2024-01-06"
---
```{r }

rm(list=base::setdiff(ls(),c('DEVMODE')))
source('util.r'); source('projects.r'); source('project.r'); source('pancancer.r'); source('optim.r')

# This rmd file is designed to run for a single optimizedparams and log file. Don't change it half-way this Rmd file. When you are running other analysis, you may need to config_reset() to switch b ack to the default files.
#config('project_optimizedparamsfile','project_optimizedparams_withcurves.yml');
#config('project_optimizedparamsfile','KeepLogs/project_optimizedparams_withcurves.yml');

if(sys_ischopserver()){ config('downloadurl.addmd5',F); }
config('project_optimizedparamsfile','project_optimizedparams_tall.yml')
sys_computername()

DEVMODE=var_get('DEVMODE',sys_issacanlap()); #this will use your defined DEVMODE; and T if not defined. Don't change the default here. Only set DEVMODE=... in the Console.

catf(list_map(list(1,2),function(...){source('util.r'); return(list(cpu=sys_cpuinfo(),ram=sys_raminfo()));},list(doparallel=T))[[1]])

res=list()

#Default parameters for projects_optimweightsandcurves()
#"OO" stands for "OPTIMIZATION OPTIONS"
OO_reset=function(){
OO=list(
  hardzeroweights='opentargetsurface_isopentargetsurface'
  ,freshstart=F #when true, we start with the defaultparams file instead of the optimizedparams file.
  ,method=csv('L-BFGS-B,SANN,Nelder-Mead,ga,Nelder-Mead,checkzeros,sfs') #the methods are listed here, but we manually use one or more below and run them separately.
  ,numrepeats=1 #For randomized methods, repeating may result in different results. How many times should we repeat?
  ,trace=1,
  methodcontrols=list(
  'SANN'=list(maxit=var_pick(DEVMODE,5,1000))
  ,'Nelder-Mead'=list(
    reltol=0.001
    ,enforcesign=T, lower=-3, upper=3 #enforcesign will trigger the use of boxconstraint version of Nelder-Mead.
  ) 
  ,sfs=list(
    abstol=0.001 #what level of accuracy improvement (forward) or reduction (reverse-elimination) are we willing to accept?
   ,suboptim=T
   ,groupweightsandcurves=F #when false, use osubtim$method=Brent to optimize individual parameters.
   ,osuboptim=list(singlepar=T,method="Brent",reltol=var_pick(DEVMODE,0.01,0.001),trace=0,enforcesign=T, lower=-3, upper=3,maxit=1000)
   #,osuboptim=list(singlepar=T,method="Nelder-Mead",reltol=var_pick(DEVMODE,0.01,0.001),trace=0,enforcesign=T,,maxit=1000)
   #,osuboptim=list(singlepar=T,method="SANN",reltol=var_pick(DEVMODE,0.1,0.001),trace=0,enforcesign=F, OO$softenforceweightsigns,maxit=100)
   #,osuboptim=list(singlepar=T,method="L-BFGS-B",trace=0,enforcesign=T, lower=-3, upper=3,maxit=100)
   ,doparallel=var_pick(DEVMODE,F,T)
   ,enforcesign=T, lower=-3, upper=3
  )
  ,parranges=list(
    abstol_accept= 0.001 #minimum improvement for updating parameters.
    ,abstol_walk= -0.001 #keep walking while allowing this much sacrifice in eval.
    ,stepsize='1%',repeatifimproved=F,maxsteps=100,enforcesign=T, lower=-3, upper=3, centerinrange=F, trace=1
  )
  ,'ga'=list(
    popSize=var_pick(DEVMODE,5,1000),maxit=var_pick(DEVMODE,5,100),stalerun=10,parallel=T
    ,enforcesign=T, lower=-3, upper=3
    ,suboptim=T,osuboptim=list(method="Nelder-Mead",poptim = 0.05,pressel=0.5,control=list(fnscale=-1,reltol=var_pick(DEVMODE,0.1,0.001),trace=0)) #it runs optim() quietly, so trace>0 would be a waste.
    ,savepopulation=T
    ,jittersuggestions=0
   )
  ,'L-BFGS-B'=list(reltol=0.01,lower=-3,upper=3)
  )
  , optimizedparamsfile=T,devmode=DEVMODE,trace=1, removeredundancy=T
  ,Iweights=NULL,Icurves=NULL #don't change these, but use subsets to control weights vs. curve optimization.
  ,subsets='weights+curves' #csv of 'weights,curves,weights+curves'. Leave NULL to decide automatically (all three on fresh start, only weights+curves on non-fresh start)
  ,doparallel=T, setupcluster=!DEVMODE||!sys_issacanlap()
  
  # after the SFS optim (STEP 1 below), we now lock in the zero-weights and zero-curves, so they do not get added again. We also lock gtex_numtissuesgt10tpm_percentile.
  ,lockzeroweights=T
  ,lockzerocurves=F
  ,lockweights='' #prevent this from changing during optimization, it usually gets smaller while others take over. We usually use/check gtex to see if something is a good target; so it is favorable to maintain a high weight for gtex.
  ,updateoptimizedparams_abstolperparam=0.001 #what is the minimum improvement for each additional feature, to update the yamlfile.
  )

#OO$method=list_remove(OO$method,'SANN');
return(OO);
}
OO=OO_reset();

#ps=pancancer_pediatricprojects(onlysurfacegenes=F,devmode=DEVMODE,doparallel=OO$doparallel,immunoadd=10);


#### Making ps only neuroblastoma to run the optimization 04.22.24 
# nbl=pancancer_project('Neuroblastoma', onlypediatric=T)
# nbl$knownpositives=c('ALK', 'GPC2', 'CD276','GFRA2', 'L1CAM', 'DLK1')
# kp=data.frame(score=c(12,12,6.5,6.5,5,5))
# rownames(kp)=c('ALK', 'GPC2', 'CD276','GFRA2', 'L1CAM', 'DLK1')
# nbl$knownpositivescores=kp
# #nbl$enrich$enrichtypes="gtex,evodevo_pediatric,compartments_sp,cirfess_spc,uniprot,depmap,opentargetsurface"
# 
# ps=projects_run(nbl)


####Making phenotype specific projects only 

# ml=pancancer_project('Melanoma', onlypediatric=T)
# ps=projects_run(ml)

# aml=pancancer_project("Acute Myeloid Leukemia", onlypediatric=T)
# ps=projects_run(aml)


# bll=pancancer_project("B-Cell Non-Hodgkin's Lymphoma", onlypediatric=T)
# ps=projects_run(bll)

# ball=pancancer_project("B-Lymphoblastic Leukemia", onlypediatric=T)
# ps=projects_run(ball)

# bl=pancancer_project("Burkitt's Lymphoma", onlypediatric=T)
# ps=projects_run(bl)

# es=pancancer_project("Ewing's Sarcoma", onlypediatric=T)
# ps=projects_run(es)

# nsclc=pancancer_project("Non-Small Cell Lung Carcinoma", onlypediatric=T)
# ps=projects_run(nsclc)

# os=pancancer_project("Osteosarcoma", onlypediatric=T)
# ps=projects_run(os)

# rms=pancancer_project("Rhabdomyosarcoma", onlypediatric=T)
# ps=projects_run(rms)

# tll=pancancer_project("T-Cell Non-Hodgkin's Lymphoma", onlypediatric=T)
# ps=projects_run(tll)

tall=pancancer_project("T-Lymphoblastic Leukemia", onlypediatric=T)
ps=projects_run(tall)


```

#History:
* Optimizing all parameters without constraints: 27% (high performance likely due to buggy rankeval where identical scores were considered inaccurately correct.)
* Excluded opentarget (hard-set at 0 to disallow its use): 12%
* Added better sign-enforcement to make sure the sign of the weights followed their expected biological interpretation and to avoid the weights from the same datasource competing with each other (some positive, some negative to compensate each other): 9%
* Changed rankeval to consider both knownpositives and knownnegatives. (Turns out we don't have any thera-knownnegatives that appear in pancancer proteins). Weighted knownpositives by their higheststatus. Ahmet uses immunoadd=10 when optimizing the weights, but you don't have to when doing pancancer_analyze().
* Changed the scaling of non-percentile-scaled features to rangemap0100.
* SFS singlepar optimization (11%).
* PENDING: Identified most promising feature from each datasource (by examining non-zero weights and sfs impact history) and limited the optimization (Iweights) to just those.


#Observations:
* Increasing ga:popSize 100->1000 did not improve the freshstart results.
* On a fresh start in ga, weights->curves order works better than curves->weights
* On a fresh start in ga, suboptim did not improve results.
* On a fresh start, SANN+Nelder performed terrible. SANN.iter=100 -> 1000 was still terrible.
* Nelder-mead abstol is not great; it causes the optimization to never start (especially with enforcesign), b/c all the first things it tries are worse. Use reltol instead.
* I no longer use SANN. It doesn't support lower/upper bounds.
* SFS.Paired weight-curve optimization did not perform well. Not used anymore.
* Using sfs suboptim L-BFGS-B is much slower (~1 hour) than Nelder-Mead (5mins) without any accuracy benefit.


```{r}
OO=OO_reset();

if(DEVMODE){ OO$methodcontrols$ga$parallel=F; OO$doparallel=F; } #Useful when debugging. Commenting this line out will revert back to T (b/c of OO_reset())
#if(DEVMODE){ OO$Iweights=1:3; OO$Icurves=1:3; }

## RUN DIFFERENT OPTIMIZATION STRATEGIES
source('util.r'); source('projects.r'); source('project.r'); source('pancancer.r'); source('optim.r'); source('optim_sfs.r')
 # cmnt='';
 # res=var_get('res'); #if there's a res, use it; otherwise use NULL.
 # OO$method='none'; cmnt=',recalculate-scores'; res=projects_optimrecipe__(res); stop();

# # # # # # # 
# #
# ### [[ 1 ]]
### SFS SINGLEPAR OPTIMIZATION: "Sequential Forward Selection with Brent's optimization to optimize the newly added parameter". Tends to limit the number of parameters selected, minimizes the computation time. A feature is added to the selected set, only if it improves the MAP score of the set by at least 0.1%.

OO$subsets='weights,curves'; OO$numrepeats=var_pick(DEVMODE,1,3); OO$method='sfs'; OO$softenforceweightsigns=F; cmnt=''; res=projects_optimrecipe__(res,lockzeroweights=F, lockzerocurves=F, updateoptimizedparams_abstolperparam=0.0015)

#OO$subsets='curves'; OO$numrepeats=var_pick(DEVMODE,1,1); OO$method='sfs'; OO$softenforceweightsigns=F; cmnt=''; res=projects_optimrecipe__(res,lockzeroweights=F, lockzerocurves=F, updateoptimizedparams_abstolperparam=0.001)
# I locked weight gtex_numtissuesgt10tpm_percentile after this step.


## [[ 2, 4, 6 ]]
### Genetic Algorithm with suboptim. Genetic algorithm with a sub-optimization of the top-scoring individual (?) in the population using Nelder-Mead. (poptim = 0.05,pressel=0.5)
#In Stage2, lockweights:gtex were used.
#OO$lockweights=csv('gtex_numtissuesgt10tpm_percentile,gtex_maxtpm_nobrain_percentile');
 OO$lockweights='';
 OO$methodcontrols$ga=modifyList(OO$methodcontrols$ga, list(optim=T, popSize=var_pick(DEVMODE,5,360), stalerun=20, loadlogpopulation=FALSE, jittersuggestions=0.02));
 OO$method='ga,Nelder-Mead';  OO$numrepeats=var_pick(DEVMODE,1,3);   cmnt=''; res=projects_optimrecipe__(res)

### [[ (3)? Optional 3rd stage ]]
## SFS selectnonzeropars (they have been well-established for inclusion). Using suboptim=multipar to identify if any additional features can be included in combination with previous ones.
# OO$lockzeroweights=F; OO$lockzerocurves=F; OO$methodcontrols$sfs$selectnonzeropars=TRUE; OO$methodcontrols$sfs$locknonzeropars=FALSE;
# OO$methodcontrols$sfs$suboptim=T; OO$methodcontrols$sfs$osuboptim=list(singlepar=FALSE,jitter=0.01,method="Nelder-Mead",reltol=var_pick(DEVMODE,0.001,0.001),enforcesign=T,maxit=1000,trace=1);
# OO$subsets='weights+curves'; OO$numrepeats=var_pick(DEVMODE,1,20); OO$method='sfs'; OO$softenforceweightsigns=F; cmnt=''; res=projects_optimrecipe__(res);


### [[ 5 ]]
## reverse-elimination
# OO$lockweights=csv('gtex_numtissuesgt10tpm_percentile,gtex_maxtpm_nobrain_percentile');
# OO$lockweights='';
# OO$methodcontrols$sfs=opt_set(OO$methodcontrols$sfs, direction='reverse', suboptim=TRUE, abstol= -0.001); 
# OO$updateoptimizedparams_abstolperparam=0.001
# OO$subsets='weights+curves'; OO$numrepeats=var_pick(DEVMODE,1,3); OO$method='sfs'; OO$softenforceweightsigns=F; cmnt=''; 
# OO$methodcontrols$sfs$osuboptim=list(singlepar=FALSE,method="Nelder-Mead",reltol=var_pick(DEVMODE,0.1,0.001),enforcesign=T,maxit=var_pick(DEVMODE,3,100),trace=1,lower=-3,upper=3,jitter=0.01); res=projects_optimrecipe__(res);
# # 
# 

### [[ Post-hoc analysis ]]
#OO$subsets='weights+curves'; OO$method='parranges'; OO$softenforceweightsigns=F; cmnt=''; OO$numrepeats=1; res=projects_optimrecipe__(res);


####################################################
### [[ 4.5 ]]
## one-by-one (NOT USED. did not really give any improvement over other optimizaitons)
#OO$methodcontrols$onebyone=opt_set(OO$methodcontrols$sfs, suboptim=T, abstol=0.001); 
#OO$updateoptimizedparams_abstolperparam=0.001
#OO$subsets='weights+curves'; OO$numrepeats=var_pick(DEVMODE,1,1); OO$method='onebyone'; OO$softenforceweightsigns=F; cmnt=''; 
#OO$methodcontrols$onebyone$osuboptim=list(singlepar=TRUE,method="Nelder-Mead",reltol=var_pick(DEVMODE,0.1,0.001),enforcesign=T,maxit=var_pick(DEVMODE,3,100),trace=1,lower=-3,upper=3,jitter=0.05); res=projects_optimrecipe__(res);


#OO$method='Nelder-Mead'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()
#OO$method='L-BFGS-B'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()


## Fix the logfile optimscore values for enforcesign. Also updates the opentarget weight to zero and does any weight-scaling. You need to manually quench the curves if they are larger than you want.
#y=project_updateoptimizedparams(res=NULL,OO); stopfif(!is.null(y$optimscore),'Remove optimscore from the yaml file before running this. The bestparam file may have been updated in the previous step with no-enforcesign, so its optimscore is not accurate for the next step where we use enforcesign.')
#OO$method='none'; OO$optimizelogfile=T; OO$removeredundancy=F; cmnt=',recalculate-scores'; projects_optimrecipe__();


#stop('end of what we want to run for now.');
#OO$method='Nelder-Mead'; cmnt=''; projects_optimrecipe__()


###  Optimize the logfile entries with enforcesign: SANN
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,50); OO$method='SANN'; OO$optimizelogfile=T; cmnt='';  projects_optimrecipe__()
#warnf('REMINDER: I recommend that you manually remove the low-performing entries after this SANN optimization.');

###  Optimize the logfile entries with enforcesign: Nelder-Mead
#OO$methodcontrols[['Nelder-Mead']]$reltol=0.001;  OO$method='Nelder-Mead'; OO$optimizelogfile=T; cmnt=''; projects_optimrecipe__()
#warnf('I recommend that you manually remove the low-performing entries after this optimization.');


### Genetic Algorithm with suboptim. Using logfile in the population.
#OO$removeredundancy=F; OO$methodcontrols$ga=modifyList(OO$methodcontrols$ga, list(optim=F, popSize=var_pick(DEVMODE,5,90), stalerun=20, loadlogpopulation=T)); OO$method='ga,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,5);   cmnt=''; projects_optimrecipe__();





#OO$method='checkzeros'; cmnt=''; statsfile=getdatafile(paste0('checkzeros_stats_noopentarget',var_pick(DEVMODE,'.devel',''),'.xlsx')); stopfif(io_isfileandnotempty(statsfile),'You must manually remove (or rename for backup) the stats file [ %s ] before running this code.',statsfile); res=projects_optimrecipe__(); data_writefile(res$stats,statsfile)
  

#Miscelleneous
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,100); OO$method='SANN,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,3);   cmnt=',notfresh,noopentarget';
#OO$methodcontrols[['Nelder-Mead']]=list(reltol=0.001); OO$method='Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget';
#OO$method='none';  OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget'; projects_optimrecipe__()


```













```{r}
stopf('SANDBOX OR MANUAL/ONE-TIME CODE SECTIONS BELOW HERE. NOT INTENDED FOR Run-All.')
```


```{r}
source('projects.r'); source('util.r')
### SFS Forward Selection feature impact summary. See the impact.summary.xlsx in each of these folders.

library(ggplot2);

for(folder in csv('sfs.forward,sfs.reverse.withoptim,sfs.reverse.withoutoptim')){
  dg=projects_sfs_summarizeimpact_frommultirun(folder=folder, grouped=T,overwrite=F)
  dg$datasource[which(dg$datasource == 'expr')]='cancer_exp'
  dg$datasource <- factor(dg$datasource, levels=dg$datasource)
  dg$impact=dg$impact*100
  g=ggplot(dg, aes(x=datasource,y=impact)) + geom_bar(stat = 'identity',color='black', fill='darkgrey') + ggtitle(folder) + theme_classic() + ylab("MAP Score Impact (%)") + theme(axis.text.x=element_text(angle = 90, hjust=0.95,vjust=0.2)) + theme(text = element_text(size=20,family = "Times New Roman"), legend.position = "none") + xlab('') + ggtitle('');
  print(g);
}

```



```{r}
### Compare different optimized yml files to study the features they use
source('util.r'); source('projects.r'); source('project.r'); source('vec_rescale.r')

projects_compareoptimizedparamsfiles();

```

