---
title: "Optimize Immunotar Parameters using Pancancer Dataset"
date: "2024-01-06"
---
```{r }
rm(list=base::setdiff(ls(),c('DEVMODE')))
source('util.r'); source('projects.r'); source('project.r'); source('pancancer.r'); source('optim.r')

# This rmd file is designed to run for a single optimizedparams and log file. Don't change it half-way this Rmd file. When you are running other analysis, you may need to config_reset() to switch b ack to the default files.
#config('project_optimizedparamsfile','project_optimizedparams_withcurves.yml');
#config('project_optimizedparamsfile','KeepLogs/project_optimizedparams_withcurves.yml');

if(sys_ischopserver()){ config('downloadurl.addmd5',F); }
sys_computername()

DEVMODE=var_get('DEVMODE',sys_issacanlap()); #this will use your defined DEVMODE; and T if not defined. Don't change the default here. Only set DEVMODE=... in the Console.

catf(list_map(list(1,2),function(...){source('util.r'); return(list(cpu=sys_cpuinfo(),ram=sys_raminfo()));},list(doparallel=T))[[1]])

res=list()

OO_reset=projects_optimoptions; #OO_reset is now an alias for projects_optimoptions(). You can use either.
OO=OO_reset();
OO$doparallel=F

ps=pancancer_pediatricprojects(onlysurfacegenes=F,devmode=DEVMODE,doparallel=OO$doparallel,immunoadd=10);
```

#History:
* Optimizing all parameters without constraints: 27% (high performance likely due to buggy rankeval where identical scores were considered inaccurately correct.)
* Excluded opentarget (hard-set at 0 to disallow its use): 12%
* Added better sign-enforcement to make sure the sign of the weights followed their expected biological interpretation and to avoid the weights from the same datasource competing with each other (some positive, some negative to compensate each other): 9%
* Changed rankeval to consider both knownpositives and knownnegatives. (Turns out we don't have any thera-knownnegatives that appear in pancancer proteins). Weighted knownpositives by their higheststatus. Ahmet uses immunoadd=10 when optimizing the weights, but you don't have to when doing pancancer_analyze().
* Changed the scaling of non-percentile-scaled features to rangemap0100.
* SFS singlepar optimization (11%).
* PENDING: Identified most promising feature from each datasource (by examining non-zero weights and sfs impact history) and limited the optimization (Iweights) to just those.


#Observations:
* Increasing ga:popSize 100->1000 did not improve the freshstart results.
* On a fresh start in ga, weights->curves order works better than curves->weights
* On a fresh start in ga, suboptim did not improve results.
* On a fresh start, SANN+Nelder performed terrible. SANN.iter=100 -> 1000 was still terrible.
* Nelder-mead abstol is not great; it causes the optimization to never start (especially with enforcesign), b/c all the first things it tries are worse. Use reltol instead.
* I no longer use SANN. It doesn't support lower/upper bounds.
* SFS.Paired weight-curve optimization did not perform well. Not used anymore.
* Using sfs suboptim L-BFGS-B is much slower (~1 hour) than Nelder-Mead (5mins) without any accuracy benefit.


```{r}
OO=OO_reset();
if(DEVMODE){ OO$methodcontrols$ga$parallel=F; OO$doparallel=F; } #Useful when debugging. Commenting this line out will revert back to T (b/c of OO_reset())
#if(DEVMODE){ OO$Iweights=1:3; OO$Icurves=1:3; }


res=projects_optim_multistage(ps,optimres=res,devmode=DEVMODE);


### [[ Post-hoc analysis ]]
#OO$subsets='weights+curves'; OO$method='parranges'; OO$softenforceweightsigns=F; cmnt=''; OO$numrepeats=1; res=projects_optimrecipe__(res);


####################################################


#OO$method='Nelder-Mead'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()
#OO$method='L-BFGS-B'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()


#stop('end of what we want to run for now.');
#OO$method='Nelder-Mead'; cmnt=''; projects_optimrecipe__()


###  Optimize the logfile entries with enforcesign: SANN
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,50); OO$method='SANN'; OO$optimizelogfile=T; cmnt='';  projects_optimrecipe__()
#warnf('REMINDER: I recommend that you manually remove the low-performing entries after this SANN optimization.');

###  Optimize the logfile entries with enforcesign: Nelder-Mead
#OO$methodcontrols[['Nelder-Mead']]$reltol=0.001;  OO$method='Nelder-Mead'; OO$optimizelogfile=T; cmnt=''; projects_optimrecipe__()
#warnf('I recommend that you manually remove the low-performing entries after this optimization.');


### Genetic Algorithm with suboptim. Using logfile in the population.
#OO$removeredundancy=F; OO$methodcontrols$ga=modifyList(OO$methodcontrols$ga, list(optim=F, popSize=var_pick(DEVMODE,5,90), stalerun=20, loadlogpopulation=T)); OO$method='ga,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,5);   cmnt=''; projects_optimrecipe__();





#OO$method='checkzeros'; cmnt=''; statsfile=getdatafile(paste0('checkzeros_stats_noopentarget',var_pick(DEVMODE,'.devel',''),'.xlsx')); stopfif(io_isfileandnotempty(statsfile),'You must manually remove (or rename for backup) the stats file [ %s ] before running this code.',statsfile); res=projects_optimrecipe__(); data_writefile(res$stats,statsfile)
  

#Miscelleneous
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,100); OO$method='SANN,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,3);   cmnt=',notfresh,noopentarget';
#OO$methodcontrols[['Nelder-Mead']]=list(reltol=0.001); OO$method='Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget';
#OO$method='none';  OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget'; projects_optimrecipe__()


```













```{r}
stopf('SANDBOX OR MANUAL/ONE-TIME CODE SECTIONS BELOW HERE. NOT INTENDED FOR Run-All.')
```


```{r}
source('projects.r'); source('util.r')
### SFS Forward Selection feature impact summary. See the impact.summary.xlsx in each of these folders.

library(ggplot2);

for(folder in csv('sfs.forward,sfs.reverse.withoptim,sfs.reverse.withoutoptim')){
  dg=projects_sfs_summarizeimpact_frommultirun(folder=folder, grouped=T,overwrite=F)
  dg$datasource[which(dg$datasource == 'expr')]='cancer_exp'
  dg$datasource <- factor(dg$datasource, levels=dg$datasource)
  g=ggplot(dg, aes(x=datasource,y=impact)) + geom_bar(stat = 'identity',color='black', fill='darkgrey') + ggtitle(folder) + theme_classic() + ylab("MAP Score Impact") + theme(axis.text.x=element_text(angle = 90, hjust=0.95,vjust=0.2)) + theme(text = element_text(size=20,family = "Times New Roman"), legend.position = "none") + xlab('') + ggtitle('');
  print(g);
}

```



```{r}
### Compare different optimized yml files to study the features they use
source('util.r'); source('projects.r'); source('project.r'); source('vec_rescale.r')

projects_compareoptimizedparamsfiles();

```

