---
title: "Optimize Immunotar Parameters using Pancancer Dataset"
output: html_document
date: "2024-01-06"
---
```{r }
rm(list=base::setdiff(ls(),c('DEVMODE')))
source('util.r'); source('projects.r'); source('project.r'); source('pancancer.r'); source('optim.r')

# This rmd file is designed to run for a single optimizedparams and log file. Don't change it half-way this Rmd file. When you are running other analysis, you may need to config_reset() to switch b ack to the default files.

config('project_optimizedparamsfile','project_optimizedparams_noopentarget_RS.02.27.24.yml');

if(sys_ischopserver()){ config('downloadurl.addmd5',F); }
sys_computername()

DEVMODE=var_get('DEVMODE',var_pick(sys_computername()=='sacanlap',T,F)); #this will use your defined DEVMODE; and T if not defined. Don't change the default here. Only set DEVMODE=... in the Console.

catf(list_map_parallel(list(1),function(...){source('util.r'); return(list(cpu=sys_cpuinfo(),ram=sys_raminfo()));})[[1]])

optimres=list()

#Default parameters for projects_optimweightsandcurves()
#"OO" stands for "OPTIMIZATION OPTIONS"
OO_reset=function(){
OO=list(
  hardzeroweights='opentargetsurface_isopentargetsurface'
  ,freshstart=F #when true, we start with the defaultparams file instead of the optimizedparams file.
  ,method=csv('L-BFGS-B,SANN,Nelder-Mead,ga,Nelder-Mead,checkzeros') #the methods are listed here, but we manually use one or more below and run them separately.
  ,numrepeats=1 #For randomized methods, repeating may result in different results. How many times should we repeat?
  ,trace=1,
  methodcontrols=list(
  'SANN'=list(maxit=var_pick(DEVMODE,5,1000))
  ,'Nelder-Mead'=list(
    reltol=0.001
    ,enforcesign=T, lower=-3, upper=3 #enforcesign will trigger the use of boxconstraint version of Nelder-Mead.
  ) 
  ,checkzeros=list(
    abstol=0.001 #what level of accuracy reduction are we willing to accept?
   ,suboptim=T,osuboptim=list(method="Nelder-Mead",reltol=var_pick(DEVMODE,0.01,0.001),trace=1)
  )
  ,'ga'=list(
    popSize=var_pick(DEVMODE,5,1000),maxit=var_pick(DEVMODE,5,100),stalerun=10,parallel=T
    ,enforcesign=T, lower=-3, upper=3
    ,suboptim=T,osuboptim=list(method="Nelder-Mead",poptim = 0.05,pressel=0.5,control=list(fnscale=-1,reltol=var_pick(DEVMODE,0.1,0.001),trace=0)) #it runs optim() quietly, so trace>0 would be a waste.
    ,savepopulation=T
   )
  ,'L-BFGS-B'=list(reltol=0.01,lower=-3,upper=3)
  )
  , doparallel=T, doclusterexport=F, optimizedparamsfile=T,devmode=DEVMODE,trace=1, removeredundancy=T
  ,Iweights=NULL,Icurves=NULL #don't change these, but use subsets to control weights vs. curve optimization.
  ,subsets='weights' #csv of 'weights,curves,weights+curves'. Leave NULL to decide automatically (all three on fresh start, only weights+curves on non-fresh start)
  )
#OO$method=list_remove(OO$method,'SANN');
return(OO);
}

OO=OO_reset();

ps=pancancer_pediatricprojects(onlysurfacegenes=F,devmode=DEVMODE,doparallel=OO$doparallel, recache=T);

projects_setupcluster(ps,doexport=T); #this takes time; you only need to do it once, but do it again if you make changes to any of the evaluation functions.
```

#Observations:
* Increasing ga:popSize 100->1000 did not improve the freshstart results.
* On a fresh start in ga, weights->curves order works better than curves->weights
* On a fresh start in ga, suboptim did not improve results.
* On a fresh start, SANN+Nelder performed terrible. SANN.iter=100 -> 1000 was still terrible.
* Nelder-mead abstol is not great; it causes the optimization to never start (especially with enforcesign), b/c all the first things it tries are worse. Use reltol instead.


```{r}
OO=OO_reset();

OO$Iweights=names(ps[[1]]$weights)[-which(names(ps[[1]]$weights) %in% c("gtex_numsamplesgt10tpm_nobrain_percentile", "evodevo_pediatric_maxrpkm_percentile", "expr_median_percentile"))] #add in the ones that I set to a defined limit
OO$Icurves=list()

OO$Iweights

ps[[1]]$weights
  
#if(DEVMODE){ OO$methodcontrols$ga$parallel=F; OO$doparallel=F; } #Useful when debugging. Commenting this line out will revert back to T (b/c of OO_reset())

## RUN DIFFERENT OPTIMIZATION STRATEGIES

source('util.r'); source('projects.r'); source('project.r'); source('pancancer.r'); source('optim.r')
cmnt='';

OO$subsets='weights';

DEVMODE=F
OO$method='Nelder-Mead'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()

ps=pancancer_pediatricprojects(onlysurfacegenes=F,devmode=DEVMODE,doparallel=OO$doparallel, recache=T);

OO$method='checkzeros'; cmnt=''; optimres=projects_optimrecipe__();
  

#OO$method='L-BFGS-B'; OO$softenforceweightsigns=F; cmnt=''; projects_optimrecipe__()


## Fix the logfile optimscore values for enforcesign. Also updates the opentarget weight to zero and does any weight-scaling. You need to manually quench the curves if they are larger than you want.
#y=project_updateoptimizedparams(res=NULL,OO); stopfif(!is.null(y$optimscore),'Remove optimscore from the yaml file before running this. The bestparam file may have been updated in the previous step with no-enforcesign, so its optimscore is not accurate for the next step where we use enforcesign.')
#OO$method='none'; OO$optimizelogfile=T; OO$removeredundancy=F; cmnt=',recalculate-scores'; projects_optimrecipe__();


#stop('end of what we want to run for now.');
#OO$method='Nelder-Mead'; cmnt=''; projects_optimrecipe__()


###  Optimize the logfile entries with enforcesign: SANN
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,50); OO$method='SANN'; OO$optimizelogfile=T; cmnt='';  projects_optimrecipe__()
#warnf('REMINDER: I recommend that you manually remove the low-performing entries after this SANN optimization.');

###  Optimize the logfile entries with enforcesign: Nelder-Mead
#OO$methodcontrols[['Nelder-Mead']]$reltol=0.001;  OO$method='Nelder-Mead'; OO$optimizelogfile=T; cmnt=''; projects_optimrecipe__()
#warnf('I recommend that you manually remove the low-performing entries after this optimization.');


### Genetic Algorithm with suboptim. Using logfile in the population.
# OO$removeredundancy=F; OO$methodcontrols$ga=modifyList(OO$methodcontrols$ga, list(optim=F, popSize=var_pick(DEVMODE,5,90), stalerun=20, loadlogpopulation=T)); OO$method='ga,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,5);   cmnt=''; projects_optimrecipe__();


### Genetic Algorithm again, this time with subobtim
#OO$methodcontrols$ga=modifyList(OO$methodcontrols$ga, list(optim=T, popSize=var_pick(DEVMODE,5,180), stalerun=20, loadlogpopulation=T)); OO$method='ga,Nelder-Mead';  OO$numrepeats=var_pick(DEVMODE,1,5);   cmnt=''; projects_optimrecipe__()

#OO$method='checkzeros'; cmnt=''; statsfile=getdatafile(paste0('checkzeros_stats_noopentarget',var_pick(DEVMODE,'.devel',''),'.xlsx')); stopfif(io_isfileandnotempty(statsfile),'You must manually remove (or rename for backup) the stats file [ %s ] before running this code.',statsfile); optimres=projects_optimrecipe__(); data_writefile(optimres$stats,statsfile)
  

#Miscelleneous
#OO$methodcontrols$SANN$maxit=var_pick(DEVMODE,5,100); OO$method='SANN,Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,3);   cmnt=',notfresh,noopentarget';
#OO$methodcontrols[['Nelder-Mead']]=list(reltol=0.001); OO$method='Nelder-Mead'; OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget';
#OO$method='none';  OO$numrepeats=var_pick(DEVMODE,1,1);   cmnt=',notfresh,noopentarget'; projects_optimrecipe__()


```















```{r}
stopf('SANDBOX CODE BELOW HERE. NOT INTENDED FOR PRODUCTION.')
```


```{r}
source('util.r'); source('projects.r'); source('project.r');

```

